---
title: "First inspection of German Microcensus Datenstrukturfile"
author: "Felix Zaussinger"
date: "31.03.2022"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 2
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE, results='hide'}

# set working directory (has to be changed at GWAP)
rootdir = "T:/Documents/Projects/04_jrc_green-skills-regional/03_data-analysis/re4gt"

# set knitr options
knitr::opts_knit$set(root.dir = rootdir)
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, results='asis')
options(digits=4)

# remove all objects from R workspace
rm(list=ls())

# Load necessary packages
library(tidyverse)
library(readxl)

# set plotting theme
# theme_set(theme_classic())

```

# Read data set  {-}

```{r, load_data, include=FALSE}

fpath_dsf_csv <- file.path(getwd(), "data", "raw", "lfs", "de", "DSF_MZ 2019.CSV")


# specify data types for select cols
colClasses=c(
  "EF951"="numeric",
  "EF952"="numeric",
  "EF953"="numeric"
  )

# read DSF, csv version
dsf <- read.csv(
  file=fpath_dsf_csv,
  header=TRUE,
  sep=";",
  dec=",",
  strip.white=TRUE,
  colClasses = colClasses
)

# replace all values < 0 with NA ("Not applicable" cases) 
dsf[dsf < 0] = NA
```

# Preprocessing of data set  {-}

## Step 1: Selection of variables for analysis {-}

Filters
  - EF39, Activity status (filter out unemployed persons)

Demographic characteristics
  - EF44: Age (demographics)
  - EF46: Sex (demographics)
  
Occupation
  - "EF114", "EF114UG1", "EF114UG2", "EF114UG3", "EF114UG4" : Current occupation, 5- to 1-digit (KldB 2010)
  
Industry
  - "EF137", "EF137UG1" : Economic activity, 3- to 2-digit (WZ 2008)
  
Region
  - EF189: Region of workplace (NUTS 2)
  - EF196: Region of workplace (NUTS 3)
  - EF564: Regionale Anpassungsschicht (> 500'000)
  - EF568: Regionale Untergruppe (> 100'000)
  - EF195: Place of work in DE?


Workplace characteristics
  - EF172: Main employment practiced (type of task)
  - EF174: Department of workplace (department of worker)
  - EF175: Status within enterprise/authority (employee/worker/official/self-employed)

Income
  - EF436: Personal net income of last month (required)
  - EF442: Avg monthly income (voluntary)
  
Education
  - EF517: highest level of educational/vocational training (ISCED)
  - EF540: very similar to EF517 (or same)
  
Hochrechnungsfaktoren
  - EF951: (bis EF564, Reg. Anp. Schichten)
  - EF952: (bis EF564, Reg. Anp. Schichten)
  - EF953: (bis EF568, Regionale Untergruppe = NUTS 3?)
  
## Step 2: coding of non-responses {-}

## Step 3: filtering by conditions {-}
labour status, age, etc.

```{r, load_data, include=FALSE}

# selection of core variables
# NOTE: uncomment commented lines when working with real data at GWAP
vars = c(
  # filtering
  "EF39",
  # demographics
  "EF44", 
  "EF46",
  # occupation
  "EF114", 
  "EF114UG1", 
  "EF114UG2", 
  "EF114UG3", 
  "EF114UG4",
  # industry
  "EF137",
  "EF137UG1", 
   # workplace characteristics
  "EF172",
  # "EF174",
  "EF175",
  # place of work
  "EF189", 
  # "EF196",
  "EF195",
  "EF564",
  # "EF568",
  # income
  "EF436",
  "EF442",
  # education
  "EF517",
  "EF540",
  "EF564",
  # projection factors
  "EF951",
  "EF952",
  "EF953"
  )

dsf_reduced <- dsf %>%
  select(all_of(vars))

# codes for non-responses (9 - 99999)
non_response_codes <- c(9, 99, 999, 9999, 99999)

# Code nan values due to non-responses (9 - 99999) for selected variables
dsf_nans_coded <- dsf_reduced %>%
  mutate(
    EF114 = replace(EF114, EF114 %in% non_response_codes, NA),
    EF114UG1 = replace(EF114UG1, EF114UG1 %in% non_response_codes, NA),
    EF114UG1 = replace(EF114UG1, EF114UG1 %in% non_response_codes, NA),
    EF114UG2 = replace(EF114UG2, EF114UG2 %in% non_response_codes, NA),
    EF114UG3 = replace(EF114UG3, EF114UG3 %in% non_response_codes, NA),
    EF137 = replace(EF137, EF137 %in% non_response_codes, NA),
    EF172 = replace(EF172, EF172 %in% non_response_codes, NA),
    # EF174 = replace(EF174, EF174 %in% non_response_codes, NA),
    EF189 = replace(EF189, EF189 %in% non_response_codes, NA),
    # EF196 = replace(EF196, EF196 %in% non_response_codes, NA),
    EF436 = replace(EF436, EF436 %in% c(90, 99), NA),
    EF442 = replace(EF442, EF442 %in% c(99), NA),
    EF517 = replace(EF517, EF517 %in% c(999), NA),
    EF540 = replace(EF540, EF442 %in% c(99), NA),
    )
  
# filtering criteria
age_min <- 0
age_max <- 95

# filter based on ILO/Eurostat conditions of active labour force
dsf_filtered <- dsf_nans_coded %>%
  filter(EF39 == 1) %>% # working against payment
  filter(EF44 >= age_min & EF44 <= age_max) %>% # age restriction
  filter(EF195 == 1) # filter out people working abroad 

# convert dtypes
dsf_final <- dsf_filtered %>%
   mutate_if(is.integer, as.factor)


```


# Analysis of data set  {-}

## Aggregation of yearly projection factors by occupation (5-digit)

```{r}
dsf_final %>%
  group_by(EF114) %>%
  summarise(
    sum_workers=sum(EF952)
  )
```

## Aggregation of yearly projection factors by industry (3-digit)

```{r}
dsf_final %>%
  group_by(EF137) %>%
  summarise(
    sum_workers=sum(EF952)
  )
```

## Aggregation of yearly projection factors by occupation (5-digit) and industry (3-digit)

```{r}
dsf_final %>%
  group_by(EF114, EF137) %>%
  summarise(
    sum_workers=sum(EF952)
  )
```

## Aggregation of yearly projection factors by occupation (5-digit) and region (reg. anp.)

```{r}
dsf_final %>%
  group_by(EF114, EF564) %>%
  summarise(
    sum_workers=sum(EF952)
  )
```